---
title: "R Notebook-Pepperwood Hobo Data Analysis"
output: html_notebook
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

R Markdown - Individual files 

```{r}
pws <-read.csv("data/PPW_HOBO_2014_Master.csv")
str(pws)
summary(pws)
```

Check the imported columns

You can also embed plots, for example:
```{r}
str(pws)
head(pws)
pws$tsfmt<- as.POSIXct(pws$Date.Time,format="%Y-%m-%d %H:%M:%S")
head(pws)
```

Lets try to do the summary by site, day for the 2014 data for RH and Temperature
```{r}
head(pws %>%mutate(HR= format(tsfmt,"%H"),DATE=format(tsfmt,"%Y-%m-%d")) %>% group_by(Plot,DATE,HR) %>% summarise(minHrlyTemp = min(Temp),maxHrlyTemp = max(Temp),meanHrlyTemp = mean(Temp),minHrlyRH = min(RH),maxHrlyRH = max(RH),meanHrlyRH = mean(RH)) %>% as.data.frame())
```

Lets write to CSV

```{r}
pws %>%mutate(HR= format(tsfmt,"%H"),DATE=format(tsfmt,"%Y-%m-%d")) %>% group_by(Plot,DATE,HR) %>% summarise(minHrlyTemp = min(Temp),maxHrlyTemp = max(Temp),meanHrlyTemp = mean(Temp),minHrlyRH = min(RH),maxHrlyRH = max(RH),meanHrlyRH = mean(RH)) %>% as.data.frame() %>% write.csv("data/Hobo-2014-Summary-Hourly.csv")
```

Now, lets create the daily file. We can follow the same, but now summarize by DATE
```{r}
pws %>%mutate(HR= format(tsfmt,"%H"),DATE=format(tsfmt,"%Y-%m-%d")) %>% group_by(Plot,DATE) %>% summarise(minHrlyTemp = min(Temp),maxHrlyTemp = max(Temp),maxHrlyTemp = max(Temp),minHrlyRH = min(RH),maxHrlyRH = max(RH),meanHrlyRH = mean(RH)) %>% as.data.frame() %>% write.csv("data/Hobo-2014-Summary-Daily.csv")
```

Lets validate our files, criteria a) Number of collection points 

```{r}
ht2014<- read.csv('./data/Hobo-2014-Summary-Hourly.csv')
str(ht2014)
head(ht2014)
```

Housekeeping , convert date from factor to a datetype, and add Month

```{r}
ht2014$dt <- as.Date(ht2014$DATE, "%Y-%m-%d")
str(ht2014)
# Check for the Sanity of data
summary(ht2014)
```

There are some corrupt values in Temperature, research the issue  
```{r}
ht2014$dt <- as.Date(ht2014$DATE, "%Y-%m-%d")
str(ht2014)
# Check for the Sanity of data
summary(ht2014)
```



Plot the number of collections by day

```{r}
ht2014 %>% mutate(site = as.factor(Plot)) %>%
group_by(site) %>%
summarise(total = n()) %>% ggplot(aes(x=site,fill=total)) + 
geom_bar() + 
theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Different way of looking at same, we expect for each plot 24 points for a day, so that makes it 24*365 (approx 8760 points for a site) . We get 2 by site every hour.
```{r}
ht2014 %>%  mutate(site = as.factor(Plot)) %>%
group_by(site) %>%
summarise(total = n()) %>% ggplot(aes(x=site,y=total)) + 
geom_point() + 
theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Issue with 1348 and 1350 collection, so have less than half the required points, so we can drop it from analysis.


Now, lets look at the unusually high temperatures from June to December
```{r}
ht2014 %>% group_by(dt)  %>% summarise(avg_temp=mean(meanHrlyTemp)) %>% ggplot(aes(dt,avg_temp)) +
geom_point() + geom_smooth(alpha=.2, size=2)
```
So, does not look good, lets look at it closer. No assumption of Temp in F or C

```{r}
p1 <- ggplot(ht2014, aes(factor(Plot), meanHrlyTemp))
p1 <- p1 + geom_boxplot(aes(fill = factor(Plot)),outlier.colour = "red", outlier.size = 3) + theme(axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none") + ggtitle("Mean temperatures across Pepperwood-2014") + xlab("Site") + 
ylab("Mean. Temperatures")+ stat_summary(fun.y=mean, geom="line", aes(group=1))  +
stat_summary(fun.y=mean, geom="point")
p1
```

Issue with 1341(?) , putting it all out of whack.

```{r}
p1 <- ggplot(ht2014 %>% filter(!(Plot %in% c(1342))), aes(factor(Plot), meanHrlyTemp))
p1 <- p1 + geom_boxplot(aes(fill = factor(Plot)),outlier.colour = "red", outlier.size = 3) + theme(axis.text.x = element_text(angle = 90, hjust = 1),legend.position="none") + ggtitle("Mean temperatures across Pepperwood") + xlab("Site") + 
ylab("Mean. Temperatures")+ stat_summary(fun.y=mean, geom="line", aes(group=1))  +
stat_summary(fun.y=mean, geom="point")
p1
```

```{r}
str(ht2014)
head(ht2014 %>% mutate(MONTH= format(dt,"%m"),DAY= format(dt,"%d")) %>% as.data.frame())
```
```{r}
str(ht2014)
head(ht2014 %>% mutate(MONTH= format(dt,"%m"),DAY= format(dt,"%d")) %>% as.data.frame())
```

Lets look at mean temperatures by month

```{r}
ht2014$site<- as.factor(ht2014$Plot)
ht2014 %>% filter(!(Plot %in% c(1342))) %>% mutate(MONTH= format(dt,"%m"),DAY= format(dt,"%d")) %>% filter(MONTH %in% c('01','02','03','04','05','06','07','08','09','10','11','12')) %>% group_by(site,MONTH,DAY) %>%
    summarise(
              min_temp = min(meanHrlyTemp),
              max_temp = max(meanHrlyTemp),
              mean_temp = max(meanHrlyTemp),
              diu_fluct = max_temp - min_temp ,
              total = n()) %>% ggplot(aes(site,mean_temp, color=MONTH)) + geom_point() +  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Mean temperatures by Month - Site n Month - 2014") + xlab("Site") + ylab("Measure of Magnitude")
```

Looks not ok, confirms the assumption that it has switced(ing) between F and C , lets look at daily Diurnal fluctuation 

```{r}
ht2014$site<- as.factor(ht2014$Plot)
ht2014 %>% filter(!(Plot %in% c(1342))) %>% mutate(MONTH= format(dt,"%m"),DAY= format(dt,"%d")) %>% filter(MONTH %in% c('01','02','03','04','05','06','07','08','09','10','11','12')) %>% group_by(site,MONTH,DAY) %>%
    summarise(
              min_temp = min(meanHrlyTemp),
              max_temp = max(meanHrlyTemp),
              diu_fluct = max_temp - min_temp ,
              total = n()) %>% ggplot(aes(site,diu_fluct, color=MONTH)) + geom_point() +  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Diurnal fluctuation - Site n Month - 2014") + xlab("Site") + ylab("Measure of Magnitude")
```